---
title: "Activity Monitoring Devices Analysis"
output: html_document
---

## Reproducible Research - Data Science Specialization

### Project I

This project makes use of data from a personal activity monitoring device . This device collects data at 5 miute intervals through out the day. The data consists of two months of data from an anonymous individual collected during de months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.


### Loading and preprocessing data

First we load the data collected from the motion device:

```{r}

mov <- read.csv("c:/activity.csv", header=TRUE, sep="," , quote = "\"")
mov$date <- as.Date(mov$date, format="%Y/%m%d")
```


### Total steps per day

We are interested in the mean total number of steps taken per day. In this case we exclude the missing values>

```{r}
hist(mov$steps, col="blue",main = "Total Steps", xlab="Steps")
```

We also calculate the mean and median of the total steps

```{r}
mean_mov <- mean(mov$steps, na.rm=TRUE)
mean_mov

median_mov <- median(mov$steps, na.rm=TRUE)
median_mov
```


###Daily Activity Pattern

We make a time series of the 5 minute interval (x-axis) and the average number of steps taken (y-axis)

```{r}
library(ggplot2)
qplot(interval,steps, data=mov )
```


###Missing Values

We calculate the missing values in the data set:
```{r}
summary(mov)
```

We are going to create a new data set replacing the values of NA in the steps column for the average 5 minute interval value

```{r}
new_mov <- mov
new_mov$steps[is.na(new_mov$steps)] <- mean_mov
summary(new_mov)
```

We calculate again the histogram, mean and median with the new data set

```{r}
hist(new_mov$steps, col="blue",main = "Total Steps", xlab="Steps")
mean_new_mov <- mean(new_mov$steps, na.rm=TRUE)
mean_new_mov

median_new_mov <- median(new_mov$steps, na.rm=TRUE)
median_new_mov
```

As observed, the replace method used has no impact on the mean, but increase the frecuency of the first range of observations


### Differences in activity patterns

We include a weekday column in the mov data set

```{r}
mov$weekday <- weekdays(mov$date)
summary(mov)
```